---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

### install and load needed libraries
```{r}
installif <- function(p) {
    if (!p %in% rownames(installed.packages()))
        install.packages(p)
    TRUE
}
sapply(c("dplyr", "ggplot2", "lubridate"), installif)
library(dplyr)
library(ggplot2)
library(lubridate)
```

### fetch the zip file from internet and extract content (if needed)

```{r}
destZipFile <- './activity.zip'
fileUrl<-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
if(!file.exists(destZipFile)) {
    download.file(fileUrl, destfile = destZipFile, method = "curl")
    unzip(destZipFile, overwrite = T)
}
```

### fix a seed to show reproducible results
```{r}
set.seed(1234)
```

### reads the content
```{r}
df <- read.csv("./activity.csv")
```

Structure of dataframe...
```{r}
str(df)
```
Some basic statistics...
```{r}
summary(df)
```

let's fix the dates and use a clean version without missing values
```{r}
df <- df %>% mutate(fixeddate = ymd(date)) %>% select(-date)
cleandf <- na.omit(df)

```

## What is mean total number of steps taken per day?
```{r}

meantotaldf <- cleandf %>% 
    group_by(fixeddate) %>% 
    summarise(totsteps=sum(steps))

as_tibble(meantotaldf)
```
Let's check how this looks in a plot, with some quick categories based on the level of activity:
```{r totstepsday}
ggplot(meantotaldf, aes( fixeddate, totsteps ) ) + 
    geom_point() + 
    geom_smooth(method=lm, color="tomato") +
    labs(title = 'Mean total steps per day', x = 'Day', y = 'Total steps') 
```

As you can see, the red line (linear model) is also showing a the mean of the total steps per day.

## What is the average daily activity pattern?

```{r avgdaypattern}

dpattern <- cleandf %>% 
    group_by(interval) %>% 
    summarise(steps=mean(steps))

ggplot(dpattern, aes( interval, steps )) + 
    geom_point(alpha=3/10) + 
    geom_smooth(color="tomato") +
#    geom_step(data=maxperday) +
    labs(title = 'Avg activity pattern', x = 'Interval', y = 'Mean steps') 

```

## Imputing missing values

How many NAs do we have?
```{r}
summary(df)
```
It looks like we have 2304 NA's. We are going to use Amelia library to complete (impute) it:
```{r}
installif("Amelia")
library(Amelia)
amelia_fit <- amelia(
    df, 
    m=1, 
    parallel = "multicore", 
    bounds = matrix( c(1,0,max(df$steps)), nrow=1, ncol=3))

imputed <- amelia_fit$imputations[[1]]

print(dim(df)==dim(imputed))
summary(imputed)
```

## Are there differences in activity patterns between weekdays and weekends?

```{r avgweekpattern}
wpattern <- imputed %>% 
    mutate( wday = weekdays(fixeddate)) %>%
    group_by(wday, interval) %>% 
    summarise(meansteps=mean(steps))

ggplot(wpattern, aes( interval, meansteps )) + 
    geom_point(alpha=3/10) + 
    geom_smooth( color="tomato") +
    facet_wrap(wday~.) +
    labs(title = 'Avg activity pattern per day of the week', x = 'Interval', y = 'Mean steps') 

```
